---
title: "Use Python in Rstudio"
author: "Georgy Makarov"
date: "5/2/2022"
output: 
  html_document:
    theme: cerulean
    highlight: arrow
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo    = TRUE)
knitr::opts_chunk$set(warning = TRUE)
knitr::opts_chunk$set(message = TRUE)
```

## Summary

`Rstudio` allows to develop and run `python` scripts. This text shows how to
prepare `Rstudio` to run `python` scripts. Preparation includes configuring 
`python`. This text does not cover installation of `python`. The following guide
applies to local `Ubuntu` machine. The text shows **2** ways: using `Linux` 
terminal and using `Rstudio` methods.

## Terminal method

At first, you work in `terminal` of your machine. In terminal go to your project
directory. Install virtual environment. Create new `python` environment for your
project. Activate your new environment. Install libraries.

```{bash, eval = FALSE}
cd <your-project-directory>
pip3 install virtualenv
virtualenv <my-env>
source <my-env>/bin/activate
pip install <lib-1> <lib-2> <lib-3>
```

Here:

 - `<your-project-directory>` -- path to your project  
 - `<my-env>` -- name of your environment  
 - `<lib-1> <lib-2> <lib-3>`-- names of libraries you want to install
 
 Then you install and configure `reticulate` library in `Rstudio`. Configure
 `Rstudio` to point to the `python` in your virtual environment. Restart `R`
 session.
 
```{r, eval = FALSE}
install.packages("reticulate")
library("reticulate")
Sys.setenv(RETICULATE_PYTHON = "<my-env>/bin/python")
```

The latter exists in all sessions. You do not need to do this command every
time.


## Rstudio method

Here you set everything up using `Rstudio` only. Install `reticulate` package.
Navigate to your project folder. Create new virtual environment and specify
packages you would like to have in it. Add path to `python` in your virtual
environment.

```{r, eval = FALSE}
install.packages("reticulate")
library("reticulate")
virtualenv_create(envname  = "<my-env>",
                  packages = c("<lib-1>", "<lib-2>", "<lib-3>"))
Sys.setenv(RETICULATE_PYTHON = "<my-env>/bin/python")
```

Activate your new virtual environment. Check that your `python` configuration
is now pointing to your virtual environment.

```{r, eval = FALSE}
use_virtualenv("<my-env>")
py_config()
```

Check that your desired packages are present in your virtual environment.

```{r, eval = FALSE}
grep(pattern = "<lib-1>|<lib-2>|<lib-3>",
     x       = as.character(py_list_packages(envname = "<my-env>")$package))
```

